\documentclass[10pt,notitlepage]{article}
\usepackage[a4paper, top=0.6in, left=0.6in, right=0.6in, bottom=0.6in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{csquotes}
\usepackage[style=ieee]{biblatex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\graphicspath{{images/}}
\addbibresource{references.bib}

\title{
  {Real or Not: How Does the Use of Natural Language Processing Improve the Classification of Disaster Tweets?}\\
  {\large Machine Learning, Group 97. Vrije Universiteit Amsterdam}
}
\author{
  Hongyu He
  \and Beata Haracewiat
  \and Matthias Debernardini
  \and Alexander Balgavy
  \and Soufiane Jounaid
}
\date{\today}

\begin{document}
  \maketitle

  \begin{abstract}
    Twitter is home to 321 million active users from all over the world.
    It is 36th on Alexa, and the verb ``tweet'' is now a mainstay in the English vernacular.
    Today, when news breaks, it breaks first on twitter.
    Often, when a disaster emerges, tweets begin to flood the Twittersphere about the situation.
    For emergency dispatchers, becoming aware of an accident as it unfolds and proactively deploying resources to the location can have a profound impact on the collateral damage.
    However, this necessitates the need for an intelligent system that can process and interpret tweets because no person or group of people can read and understand the constant stream of tweets.
    Because these tweets are mainly composed of natural language, the intelligent system needs to process the natural language so that it can output actionable suggestions.
  \end{abstract}

  \input{introduction}
  \input{method}
  \input{experiments}
  \input{results-and-evaluation}
  \input{conclusion}
  \input{discussion}

  \printbibliography

  \input{appendix}
\end{document}
